---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r include = F}
library(dplyr)
library(data.table)
library(gdsfmt)
library(ggplot2)
library(mnda)

source("~/kul/dex-stim-human-array-isns/code/00_functions/load_gds_files.R")
source("~/kul/dex-stim-human-array-isns/code/00_functions/get_chord_diagramm.R")

cbPalette <- c("grey", "#0072B2")
```

```{r include = F}
rslt_dir <- "/binder/mgp/workspace/2020_DexStim_Array_Human/dex-stim-human-isns/results/5_fold_cv/all/"
setwd(rslt_dir)
```

```{r load-data, include = F}
treatment <- "dex"
dex_global_net <- readRDS(paste0(rslt_dir, treatment, "/networks_pearson_cor/pearson_global_chr_all.rds"))

treatment <- "veh"
veh_global_net <- readRDS(paste0(rslt_dir, treatment, "/networks_pearson_cor/pearson_global_chr_all.rds"))
```

# Overlap between dex and veh features

```{r}
dex_features <- rownames(dex_global_net$network) # 5'388
veh_features <- rownames(veh_global_net$network) # 4'519

olap <- intersect(dex_features, veh_features) # 758
olap_cpgs <- olap[grepl("cg", olap, ignore.case = T)] # 566
olap_snps <- olap[grepl("rs", olap, ignore.case = T)] # 192
```

# Filter significant associations

```{r include = F}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  if (is.null(pmat)){
    data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut])
  } else{
    data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]) 
  }
}
```

```{r}
dex_global_net_flatten <- flattenCorrMatrix(dex_global_net$network, dex_global_net$pvalues) %>% 
  mutate(fdr = p.adjust(p, method = "BH")) %>% setDT()
veh_global_net_flatten <- flattenCorrMatrix(veh_global_net$network, veh_global_net$pvalues) %>% 
  mutate(fdr = p.adjust(p, method = "BH")) %>% setDT()
```

```{r}
dex_global_net_flatten_sign <- dex_global_net_flatten[fdr <= 0.05]
veh_global_net_flatten_sign <- veh_global_net_flatten[fdr <= 0.05]
```


```{r load-cpg-coordintaes, include = F}
setwd("/binder/mgp/workspace/2020_DexStim_Array_Human/dex-stim-human-isns/")

treatment   <- "veh"
dnam_gds_fn <- paste0("input/dnam/mad_filtered/gds/methyl_beta_mtrx_corrected_for_cov_mad80_filtered_", treatment, ".gds")
cpg_loc     <- LoadMethylCoordinates(dnam_gds_fn) %>% data.frame() %>% setDT()
```

```{r load-snp-coordintaes, include = F}
setwd("/binder/mgp/workspace/2020_DexStim_Array_Human/dex-stim-human-isns/")

snps_gds_fn <- paste0("input/snps/ld_pruned/gds/dex_geno_imputed_maf_ld_pruned_from_gen.gds")
snp_loc     <- LoadGenotypeCoordinates(snps_gds_fn) %>% data.frame() %>% setDT()
```

```{r create-tbl-for-plt, include = F}
dex_net_cpg_loc <- cpg_loc[CpG_ID %in% union(dex_global_net_flatten_sign$row, dex_global_net_flatten_sign$column)] %>% 
  mutate(ID = CpG_ID, Type = "CpG", treatment = "GR-stimulated") %>% dplyr::select(-c(CpG_ID))
dex_net_snp_loc <- snp_loc[SNP %in% union(dex_global_net_flatten_sign$row, dex_global_net_flatten_sign$column)] %>% 
  mutate(ID = SNP, Type = "SNP", treatment = "GR-stimulated") %>% dplyr::select(-SNP)

veh_net_cpg_loc <- cpg_loc[CpG_ID %in% union(veh_global_net_flatten_sign$row, veh_global_net_flatten_sign$column)] %>% 
  mutate(ID = CpG_ID, Type = "CpG", treatment = "Baseline") %>% dplyr::select(-c(CpG_ID))
veh_net_snp_loc <- snp_loc[SNP %in% union(veh_global_net_flatten_sign$row, veh_global_net_flatten_sign$column)] %>% 
  mutate(ID = SNP, Type = "SNP", treatment = "Baseline") %>% dplyr::select(-SNP)

net_loc     <- rbind(dex_net_cpg_loc, dex_net_snp_loc, veh_net_cpg_loc, veh_net_snp_loc)
```

# Distribution of features, CpGs and SNPs across chromosmes and time points (not overlap)

```{r}
net_loc_count <- net_loc %>%
  group_by(chr, Type, treatment) %>%
  tally() %>%
  group_by(treatment) %>%
  mutate(freq = n / sum(n)) %>% 
  mutate(chr = factor(chr, levels = 1:22)) %>% setDT()
```

```{r}
ggplot(net_loc_count, aes(x = treatment, y = freq, fill = Type)) + 
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = net_loc_count[freq > 0.01], 
            aes(label = scales::percent(freq, accuracy = 1), y = freq), 
            stat = "identity", position = position_stack(vjust = 0.5), size = 3,
            color = "black") +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(.~chr, nrow = 1, strip.position = "top") +
  labs(title = "",
       x = "",
       y = "Variant frequency") +
  theme(legend.title = element_blank(), 
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 8),
        axis.title = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 0.5, colour = "black"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_fill_brewer(palette = "Set2")
  # scale_color_manual(values = alpha(cbPalette, 0.75)) +
  # scale_fill_manual(values = alpha(cbPalette, 0.75))
```

# Nr of CpGs per SNP 
```{r veh-snp-cpg_cnt, include = F}
# Take only QTLs
thrsh <- 0

veh_qtl_net <- rbind(veh_global_net_flatten_sign[row %like% "cg"][column %like% "rs"],
                     veh_global_net_flatten_sign[row %like% "rs"][column %like% "cg"]) %>%
  unique() %>%
  dplyr::select(CpG_ID = row, SNP = column, value = cor, fdr)

veh_qtl_cnt <- veh_qtl_net[abs(value) >= thrsh, ] %>% 
  group_by(SNP) %>%
  tally() %>% 
  mutate(treatment = "Baseline")
```

```{r dex-snp-cpg_cnt, include = F}
# Take only QTLs
dex_qtl_net <- rbind(dex_global_net_flatten_sign[row %like% "cg"][column %like% "rs"],
                     dex_global_net_flatten_sign[row %like% "rs"][column %like% "cg"] ) %>%
  unique() %>%
  dplyr::select(CpG_ID = row, SNP = column, value = cor, fdr)

dex_qtl_cnt <- dex_qtl_net[abs(value) >= thrsh, ] %>% 
  group_by(SNP) %>%
  tally() %>% 
  mutate(treatment = "GR-stimulated")
```

```{r}
qtl_cnt <- rbind(veh_qtl_cnt, dex_qtl_cnt)

plt.size <- 12
ggplot(qtl_cnt, aes(x = treatment, y = n, color = treatment, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  theme(legend.position = "bottom", # c(0.9, 0.9), 
        legend.title = element_blank(), 
        legend.text = element_text(size = plt.size),
        panel.grid.major = element_blank(), 
        panel.background = element_blank(),
        plot.title = element_text(size = plt.size), 
        axis.title = element_text(size = plt.size),
        axis.text = element_text(size = plt.size, colour = "black")) + 
  labs(title = "",
       x = "", y = "Nr. of CpGs per SNP") +
  scale_color_manual(values = cbPalette) +
  scale_fill_manual(values = alpha(cbPalette, 0.25))
```

# Chord plot of meQTLs

## Basline

```{r}
snp_max <- veh_qtl_cnt[order(veh_qtl_cnt$n), ][1:10,]
CustomChordDiagram(veh_global_net$network, cpg_loc, snp_loc[SNP %in% snp_max$SNP,], thrsh = 0)
```


## GR-stimulated

```{r}
CustomChordDiagram(dex_global_net$network, cpg_loc, snp_loc, thrsh = 0.0)
```

####------------
# 4. Check how many cis / trans

# modules_features <- features[do.call(c, modules)]
# 
# cpg_loc[CpG_ID %in% modules_features]
# snp_loc[SNP %in% modules_features]

####------------

# 5. Coord plot 

## 5.1. Baseline

net <- veh_global_net %>% as.matrix() %>% as.data.frame()
net <- veh_global_net[!rownames(veh_global_net) %in% olap,
                      !colnames(veh_global_net) %in% olap] %>% as.matrix() %>% as.data.frame()
CustomChordDiagram(net, cpg_loc, snp_loc, thrsh = 0.80)

## 5.2. Post-dexamethasone

net <- dex_global_net %>% as.matrix() %>% as.data.frame()
CustomChordDiagram(net, cpg_loc, snp_loc, thrsh = 0.25)

## 5.3. Net with olap features betweeb pre- and post-dex

### 5.3.1 Baseline

net <- veh_global_net[olap, olap] %>% as.matrix() %>% as.data.frame()
CustomChordDiagram(net, cpg_loc, snp_loc, thrsh = 0)





# Get CpGs' and SNPs' locations for only features present in the network "net"
net_cpg_loc <- cpg_loc[CpG_ID %in% colnames(net)] %>% mutate(ID = CpG_ID) %>% select(-CpG_ID)
net_snp_loc <- snp_loc[SNP %in% colnames(net)] %>% mutate(ID = SNP) %>% select(-SNP)
net_loc     <- rbind(net_cpg_loc, net_snp_loc)[match(rownames(net), ID)]
